\section{Initial Gene Finding Results} 

Counts of genes predicted by Braker2, GeneMark and RefSeq are shown in
table \ref{table:gene-counts}. Immediately we see that Braker2
predicts far fewer genes in all assemblies, except in the case of
\textit{Trichoderma reesei.} This is possibly due to the effects of
training the Braker2 gene model using data from \textit{Trichoderma
  reesei}, which has a significantly smaller genome in comparison to
other \textit{Trichoderma} assemblies, although genome size is not
always indicative of gene content. Regardless, there is a significant
difference in the number of genes predicted by Braker2 in comparison
to GeneMark and RefSeq. The number of genes predicted by GeneMark and
RefSeq are similar, except in the case of \textit{T. harzianum}, in
which RefSeq predicts roughly 17\% more genes than GeneMark. Braker2
consistently predicts more transcripts than GeneMark and
RefSeq. RefSeq also appears to predict multiple transcripts for each
gene but in fewer numbers than Braker2. Transcript prediction counts
in \textit{T. harzianum} from RefSeq are also interesting, with RefSeq
predicting fewer transcripts than genes. Why this is occurs is unknown
but may warrant further investigation.

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Assembly & Braker2 & & GeneMark & & RefSeq & \\ \hline
     & Genes & Transcripts & Genes & Transcripts & Genes & Transcripts \\ \hline
    DC1 & 8546 & 8637 & 11353 & 11353 & N/A & N/A \\ \hline
    Tsth20 & 8784 & 8858 & 12362 & 12362 & N/A & N/A \\ \hline
    \textit{T. reesei} & 9659 & 10175 & 9196 & 9196 & 9109 & 9118 \\ \hline
    \textit{T. harzianum} & 8314 & 8385 & 12164 & 12164 & 14269 & 14090 \\ \hline
    \textit{T. virens} & 7801 & 7863 & 11866 & 11866 & 12405 & 12406 \\ \hline
  \end{tabular}
  \caption[Gene prediction counts]{Number of genes predicted by each
    gene finder for each \textit{Trichoderma} genome.}
  \label{table:gene-counts}
\end{table}


