\section{Profiling Gene Finding Tools}

While Braker2, GeneMark and RefSeq all provide lists of possible genes
for a provided reference, the implementation of each tool is
different, requiring more or less effort to install and run depending
on the tool used. The computing platform used in this research is
hosted and managed by University of Saskatchewan services, which is
modelled around the HPC platform used by the Digital Research Alliance
of Canada and software is managed similarily. 

First, we will briefly discuss the RefSeq annotation process. RefSeq
annotation is only applied to data that is submitted to NCBI. The
RefSeq Eukaryotic Genome Annotation Pipeline\cite{NCBI2024} is a
genome annotation process developed and maintained by NCBI. The
pipeline is not directly publicly available to public users, and
requires submission of data to NCBI. Once data is submitted to NCBI,
the RefSeq annotation pipeline may be applied upon request only if the
genome is the highest quality assembly for the species in question or
if the genome is of significant interest to the scientific community,
limiting reach of the annotation process to many users. The pipeline
supplies existing RNAseq, CDS and protein sequences to NCBI's in-house
gene prediction tool Gnomon, which produces trained models for gene
prediction. While the tools used for alignment and processing of
supporting sequence information are listed, the inner workings of
Gnomon are not well documented, at least from the public perspective,
and I was unable to find Gnomon in any compilable or executable form
during my search. Recreation of the RefSeq pipeline would prove
extremely challenging if not impossible without supporting
information. Run times for this pipeline are difficult to determine
due to the hidden nature of the pipeline, unknown compute resources
and varying quantities of data used. The RefSeq annotation process
produces comprehensive outputs, including CDS sequences, translated
CDS, RNA from genomic sequences, proteins, feature counts and tables,
and finally GFF and GTF formatted annotation files for these features.

Next we will discuss the handling, installation, and running of both
Braker2 and GeneMark packages. Several points will be discussed for
each tool, with those points being sourcing and downloading,
components and prerequisites, installation, execution, and
output. GeneMark\cite{10.1093/nar/gki937} is a gene finding tool
developed by the Georgia Institute of Technology with packages
prepared for Linux and MacOS. It is provided as licensed product in
the form of a package which can be downloaded from their website after
submitting a form. Once the necessary information is submitted, the
user is provided with a key that must be placed in the appropriate
location once the software is downloaded and unpacked. The core
controlling methods of GeneMark are written in Perl, accompanied by
several Python scripts and compiled exectuables. GeneMark was tested
by the developers with Perl version 5.10, and Python 3.3. A number of
Perl dependencies are also required, which can be installed via
YAML. The user will have to know which implementation of GeneMark they
are wanting to use for their application, as GeneMark has several
variations it can run depending on the desired application. In this
work, the GeneMark-ES variant of GeneMark was executed, as it is the
self-training \textit{ab initio} GeneMark method for eukaryotic
organisms .Options required by GeneMark at runtime are documented in
the help message, and simple enough that any user with familiarity of
bioinformatics tools should be able to run GeneMark, although
documentation for use is only provided by the help message when
running the program and not online. In regards to run-time, running
the GeneMark-ES pipeline on DC1 with 56 threads finished in 16
minutes. Upon completion, GeneMark produces a GTF or GFF file of
predicted genes as well as a number of other outputs related to the
run.

Braker2\cite{Bruna2021} is hosted on GitHub as a repository that
receives relatively frequent updates with Braker3 being released while
working on this thesis. Braker is maintained by Katharina Hoff from
the University of Greifswald and is available under the Open Source
Artistic License. Installation of the repository is a straightforward
pull from GitHub. As with GeneMark, Braker2 uses a combination of
Perl, Python and other exectuables in its regular use. Downloading the
repository itself is not enough for execution, as Braker2 relies on a
number of dependencies and bioinformatics tools including Perl and
(Perl dependencies), Augustus, BamTools, BedTools, GeneMark,
StringTie, GFFRead and a few others. Manual installation of these
dependencies would be difficult, time consuming and in general advised
against. In this case, many of Braker2's requirements are satisfied by
modules already included in the environment, making installation
relatively simple if you know the ins and outs of the Digital Research
Alliance of Canada's software stack. This case still required
installation of some Perl modules in addition to loading necessary
modules. Alternatively, one could use a package manager like Anaconda
or Minoconda to handle installation of packages. This is perfectly
reasonable, but also requires knowledge specific to Anaconda, which
can be complicated and frustrating for users with little software
management experience. The newly released Braker3, also includes a
containerized version of the software, allowing users to build and
execute Braker3 with ease in its own environment. Once installation is
finished, the Barker2 pipeline is relatively straightforwad to run as
well, with excellent documentation included both online and through
the built in help message. In this research, the Braker2 pipeline was
run in two modes. The first mode is a training mode, where sequence
files are supplied to Braker2 to train a gene-calling model. The
training and gene finding steps are run as part of the same command,
and a model built using the training data is saved for bookkeeping and
future use. The training was performed using roughly 145 million
Illumina paired end RNAseq reads on the \textit{Trichoderma reesei}
genome. The training, including RNAseq alignments, and gene finding
pipeline in this case took 1 hour and 17 minutes using 60
threads. Applying the Braker2 trained gene finding model to DC1 with
60 threads took 21 minutes to complete. Run times will of course vary
depending on processing power available to the end user, but in the
case of \textit{Trichoderma} genomes, users can expect quick results
with relatively little computing power. Once annotation is complete,
Braker2 produces a GFF file containing predicted genes along with CDS
sequences and amino acid sequences their protein products. Braker2,
when trained with RNAseq data, includes an option to predict UTR
regions, but it is experimental and is not performed by default and is
not included in these runs.

In summation, Braker2 and GeneMark are not direct plug-and-play
software packages. Users should expect to encounter issues when
getting these programs running in addition to normal downloading and
unpacking of software packages so some expertise is
recommended. Neither Braker2 nor GeneMark require users to compile
software, however Braker2's dependencies may require additional
compilation and attention. Once installed, both tools are relatively
simple to use with documention available for both on the commandline
and excellent documentation available for Braker2 on their GitHub
page. Outputs from both tools are similar although Braker2 has the
ability to output coding and amino acid sequences for downstream
processing. Both tools run in reasonable amounts of time, where in the
case of smaller genomes such as \textit{Trichoderma}, users can expect
results within a few hours to a day depending on number of computing
cycles available to them.
