\documentclass[12pt]{article}
\usepackage[noblocks]{authblk}

\title{Project Proposal for Analysis of Novel \textit{Trichoderma} Genomes}
\author{Connor Burbridge}
\affil{USask NSID: cbe453 \\
  USask ID no.\ 11162928 \\
  Supervisors: Dave Schneider \& Tony Kusalik\\}

\begin{document}
\parindent=14pt
\maketitle

\clearpage
\tableofcontents
\clearpage

\section{Background}

\subsection{Trichoderma}

Crop resistance to environmental stressors is a necessity for crop
health and overall crop yields. Current popular methods for crop
protection involve the use of pesticides and genetically modified
organisms, which can be expensive and potentially politically dividing
in the case of GMOs (citation needed). In addition, crops will suffer
when soils are not sufficient for crop growth and health. Soil
insufficiencies can result in drought stress as well as nutrient
stress, leading to poor overall yields.

\textit{Trichoderma} is a type of fungi that can colonize the roots of
plants in a non-toxic, non-lethal, opportunistic symbiotic
relationship\cite{Trichoderma}. Many strains of \textit{Trichoderma}
have been shown to provide resistance to bacteria and other fungi in
soils through the use of polyketides, non-ribosomal peptide
synthetases and other antibiotic
products\cite{Trichoderma}\cite{Secretome}. Recently, two strains of
\textit{Trichoderma} have been identified in the prairie regions of
Alberta and Saskatchewan. These two strains, named Tsth20 and DC1,
have been found to have beneficial properties when used as an
inoculant for plants in the soils mentioned before. In addition to
these beneficial properties, the two strains mentioned previously
provide even further protection for plants in dry, salty soils and one
strain also has potential for use as a bioremediation tool in soils
contaminated with hydrocarbon content. However, little is known about
the mechanisms at work in these new strains, so DC1 and Tsth20 were
sequenced by the Global Institute for Food Security (no publication
yet so no reference?) in an initial attempt to better understand the
details of these genomes and secretomes.

\subsection{Genome Assembly}

Sequence assembly has been a long-standing application problem in the
field of bioinformatics\cite{assembly}. Determining the correct order
and combination of smaller subsequences into an accurate sequence
assembly is also computationally difficult in terms of compute
resources such as memory, CPU cycles and storage required for input
sequences\cite{assembly}. In addition to these difficulties, there can
be other issues encountered during asssembly due to the nature of
the data or genomes themselves, such as low quality base calls for
long read data or the inherent content of genomes themselves with
repetitive regions as an example. Insufficient data used in an
assembly may result in short, fragmented assemblies, depending on the
size of the genomes, while sequence data that is not long enough can
fail to fully capture repetitive regions in an assembly. To solve this
problem, a wide range of assembly tools have been developed with their
own unique approaches to genome assembly problems, so it is important
to use an appropriate assembler for the task at hand, and important to
evaluate the assembly thoroughly. One approach to aid in the
previously mentioned issue of assembly correctness, is to use a
combination of long and short reads in what is known as a hybrid
assembly. Combining both highly accurate short reads with deep
coverage along with less accurate but much longer reads can produce
high quality genome assemblies that capture long repetitive
regions. Assemblies must also be evaluated with measures such as N50,
L50, coverage, average contig length and total assembled length to
ensure that the genomes are well assembled, at least based on these
metrics\cite{assembly}. Following appropriate assembly protocols is
essential to the further success of a project as downstream processing
such as annotation depends on a high-quality assembly.

\subsection{Repeat Identification/Masking and Identification of AT-rich Genomic regions}
Repeat identification within assembled genomes is a problem that
needs to be considered during the genome annotation process. Regions
with long repeats can have a significant impact on genome assembly as
well as gene finding due to the limitation of short reads used in some
assemblies\cite{Repeats}. Short reads may be unable to bridge or cover
entire repeat regions within a genome, so it is important to consider
the use of long reads such as Nanopore or PacBio to provide a complete
picture of these regions when pursuing a new genome assembly
project. It is also possible for these repetitive regions to contain
genes as well, making for an interesting investigation in regards to
\textit{Trichoderma}, as fungal genomes have been shown to contain
many repeat regions with a high concentration of A and T
nucleotides\cite{fungalrepeats}. Once these repetitive regions have
been identified (by any number of existing programs), the genome could
be masked to exlude these regions in downstream processing if desired,
as these regions may be poorly assembled and may result in found genes
that may not truly exist in those regions. However, this may not be as
common today\cite{dontMask}. This may affect the gene finding process
described later and may be an interesting topic to look into
considering the large number of available gene finding programs
available.

\subsection{Gene Finding}
Gene finding is another long standing computational problem in
bioinformatics, which concerns itself with identifying potential genes
within genomes based on patterns or evidence considered by the gene
finding program. There are two common methods for gene finding, those
methods being \textit{ab initio} methods where programs search for
patterns and gene structures, while similarity or evidence-based
searches use prior information such as RNAseq data, expressed sequence
tags and protein sequences to identify genes within a new
genome\cite{GeneFinding}. Complicating the process more is the
introduction of introns and alternative splicing in eukaryotes, making
it possible for one gene to have several possible transcripts at the
same locus. Examples of \textit{ab initio} methods include tools such
as GeneMark-ES\cite{GeneMarkES} and GlimmerHMM\cite{Glimmer}, while
evidence based methods include tools such as Braker2\cite{Braker2},
which can incorporate existing data such as RNAseq, protein sequences,
etc. in gene prediction models.

As mentioned in the repeat finding section of this proposal, long
repeats and transposable elements can cause issues for gene finding
programs, making this an interesting area of research, at least for
fungal genomes such as \textit{Trichoderma}. Fungal genomes are also
interesting in the topic of gene finding as there are few programs
targeted directly at gene finding in fungal genomes while organisms
such as human, mouse, and \textit{Arabidopsis} benefit from having
many tools tested on them as they are model organisms in the field, at
least according to table 1 in (Wang, Z \textit{et
  al})\cite{GeneFinding}. How these different methods and tools
perform when applied to fungal genomes is an important consideration
as fungi have features that can benefit plant growth as mentioned
earlier.

There are also other aspects of gene finding tools that are important
to consider. These include features such as whether or not the gene
finders find non-coding RNAs, annotation of 5' and 3' UTR regions, and
in the case of ab-initio methods, the assumptions made by the
underlying models used for gene finding. These features and others,
can influence a users decision on which gene finding tool to consider
and will complicate comparative analysis of multiple gene finding
tools.


\section{Research problem}
Genome annotation, in the case of gene finding in this work, has been
a popular computational problem for decades. The identification of
possible genes provides other researchers with a valuable resource for
future research avenues. Due to the importance and popularity of this
topic, there have been a variety of gene finding tools developed, each
with their own implementation for certain applications or use
cases. The abundance of gene finding tools raises questions such as
which tool is better? Do different tools provide drastically different
results? Are there differences in the number and/or of accuracy
predicted genes in repetitive genomic regions between \textit{ab
  initio} methods and similarity-based methods?  Comparative studies
of gene finding tools have been performed before, however most of
these studies are in reference to popular eukaryotic subjects of study
such as humans, mice and \textit{Arabidopsis}. Few if any comparative
studies have been performed in fungal genomes, and even fewer in the
case of \textit{Trichoderma}.

This lack of comparative analysis in fungal genomes provides a
research avenue to compare results from the application of different
gene finding tools to new and existing \textit{Trichoderma}
assemblies. Results from using different annotation tools will allow
us to identify a core genome along with genes that are not predicted
by all tools in all genomes, or outliers. This analysis will also
allow us to compare results from both evidence-based gene finding
methods as well as \textit{ab initio} gene finding methods. One
interesting area to investigate between these methods would be in
regions of low nuclotide diversity, or the AT-rich regions common to
\textit{Trichoderma} and other fungi. It is possible that gene finding
programs may have difficulty with these regions of genomes, as they
may contain repeat regions and as mentioned before, have low
nucleotide diversity, making it unlikely but not impossible for genes
to be found in these regions.

As mentioned previously, it is important to understand the features of
these gene finders prior to developing an approach for comparing their
outputs. With this in mind, the tools selected for this research
should cover both \textit{ab initio} and evidence-based methods as
well as the identification of other features such as small RNAs and
untranslated regions as it is possible that some gene finders may
identify features such as small RNAs without explicitly stating that
they do. This information may also be useful for GIFS researchers
working with this data in the future. To cover several of these areas
of interest, the following tools are proposed for this project. Note
that other tools are also available for consideration and this set
does not have to be final.

\begin{itemize}

\item ReatMasker\cite{RepeatMasker}: a tool used to identify
  repetititve regions within a genomic sequence.
\item Infernal/Rfam\cite{Infernal}: a tool used in combination with a database to
  analyze genomic sequences for non-coding RNAs. Useful to compare
  against gene callsfrom the tools proposed below.
\item GeneMark-ES\cite{GeneMarkES}: an \textit{ab initio} HMM-based
  gene finding tool with an option for specific to fungal genomes,
\item GenomeThreader\cite{GenomeThreader}: an evidence-based gene finding
  tool that uses dynamic programming to efficiently identify splice
  candidates for genes using splice-aware RNASeq alignments as input.
\item Braker2: a gene finding tool that uses a hybrid approach
  combining information from \textit{ab initio} and evidence-based
  methods. In addition, Braker2 also has the ability to identify UTR
  regions as well

\end{itemize}

A comparative approach or pipeline will then be developed to compare
outputs and performance of these gene finding tools. Possible points
of comparison include true positives as compared to aligned sequence
data, identification of genes in repetitive regions and identification
of potential small RNAs. Results from this work will provide insight
into the use of different gene finding tools in the context of
\textit{Trichoderma}. This work will also provide a large set of
called genes spanning at minimum two newly assembled
\textit{Trichoderma} genomes, those being DC1 and Tsth20, as well as
other previously assembled \textit{Trichoderma} genomes from NCBI or
other sources.

\section{Deliverables}
\begin{itemize}
\item Lists of genes for each \textit{Trichoderma} assembly and gene
  finding tool combination considered
\item A consenus or core gene set shared by all genomes considered
  based on different gene finding tools
\item Repeat regions identified in assembled \textit{Trichoderma}
  genomes.
\item A potential list of true positives based on genes which have
  supporting RNAseq eveidence
\end{itemize}

\section{Timeline}
\begin{itemize}
\item Initial genome assemblies: Collection of existing assemblies (1 week), Assembly of DC1/Tsth20 (2-4 weeks)
\item Gene finding and annotation results: Identification of gene finding tools (1-2 weeks), Gene finding analysis on selected genomes (1-2 months)
\item Downstream analysis of gene finding results (1-2 months)
\end{itemize}

\section{References}

\begin{thebibliography}{99}
\bibitem{Trichoderma} Hermosa, R., Viterbo, A., Chet, I., Monte,
  E. (2012). Plant-beneficial effects of \textit{Trichoderma} and of
  its genes. \textit{Microbiology}. \textit{38},
  17-25. 

\bibitem{Secretome} Ramirez-Valdespino, C., Casas-Flores, S.,
  Olmedo-Monfil, V. (2019). \textit{Trichoderma} as a Model to Study
  Effector-Like Molecules. \textit{Frontiers in Microbiology} \textit{15}.

\bibitem{Braker2} Hoff, K. J., Lomsadze, A., Borodovsky, M., 
  Stanke, M. (2019). Whole-Genome Annotation with
  BRAKER. \textit{Methods Mol Biol.}, \textit{1962}, 65-95.

\bibitem{GeneMarkES} Borodovsky, M., Lomsadze, A. (2011). Eukaryotic
  Gene Prodeiction using GeneMark.hmm-E and
  GeneMark-ES. \textit{Current Protocols in
    Bioinformatics}. \textit{4}.

\bibitem{Glimmer} Majoros, W.H., Pertea, M., Salzberg, S.L. (2004) TigrScan and
  GlimmerHMM: two open-source \textit{ab initio} eukaryotic
  gene-finders. \textit{Bioinformatics}. \textit{20}(9), 2878-2879.
  
\bibitem{AUGUSTUS} Stanke, M., Morgenstern, B. (2005). AUGUSTUS: a web
  server for gene prediction in eukaryotes that allows user-defined
  constraints. \textit{Nucleic Acids Research}. \textit{33}, 465-467.
  
\bibitem{InterProScan} Jones, P., Binns, D., Chang, H. Y., Fraser,
  W., Li, W., ... Hunter, S. (2014). InterProScan 5: genome-scale
  protein function
  classification. \textit{Bioinformatics}. \textit{30}(9),
  1236-1240.

\bibitem{AnnotationSummary} Yandell, M., Ence, D. (2012). A beginner's
  guide to eukaryotic genome annotation. \textit{Nature Reviews
    Genetics}. \textit{13}, 329-342.
  
\bibitem{assembly} Sohn, J., Nam, J. (2016) The present and future of
  \textit{de novo} whole-genome assembly. \textit{Briefings in
    Bioinformatics}. \textit{19}1, 23-40.

\bibitem{Repeats} Lerat., E. (2009). Identifying repeats and
  transposable elements in sequences genomes: how to find your way
  through the dense forest of programs. \textit{Nature
    Heredity}. \textit{104}, 520-533.

\bibitem{fungalrepeats} Li, WC., Huang, CH., Chen, CL. et
  al. (2017). Trichoderma reesei complete genome sequence,
  repeat-induced point mutation, and partitioning of CAZyme gene
  clusters. \textit{Biotechnol Biofuels}. \textit{10}, 170.

\bibitem{dontMask} Slotkin, R.K. (2018). The case for not masking away
  repetitive DNA. \textit{Mobile DNA}. \textit{9}(1), 15.
  
\bibitem{GeneFinding} Wang, Z., Chen, Yazhu., Li, Y. (2016). A Brief
  Review of Computational Gene Prediction Methods. \textit{Genomics
    Proteomics Bioinformatics}. \textit{2}4, 216-221.

\bibitem{RepeatMasker} Smit, A., Hubley, R., Green, P. (2013-2015)
  RepeatMasker Open-4.0. https://www.repeeatmasker.org

\bibitem{Infernal} Nawrocki, E.P., Eddy, S.R. (2013). Infernal 1.1:
  100-fold faster RNA homology
  searches. \textit{Bioinformatics}. \textit{29}, 2933-2935.

\bibitem{GenomeThreader} Gremme, G., Brendel, V., Sparks, M.E., Kurtz,
  S. (2005) Engineering a software tool for gene structure prediction
  in higher organisms. \textit{Information and Software
    Technology}. \textit{47}, 965-978.
  
\end{thebibliography}

\end{document}
