\section{Analysis of Results}

After completion of the processing portion of this work, the results
must be processed in a useful way, which includes both the biological
implications of the gene calls as well as the computational, or gene
finding features, of the the selected programs. To better understand
how gene finders perform in these two classes, we must define an
appropriate plan for analysis of the results produced so
far. Currently, downstream analysis plan has been broken down into
several sections.

\subsection{Basic Analysis}

Basic analysis of gene finding results is an important part of this
research. Total gene, transcript and protein counts will be identified
for each genome and gene finding tool combination. Comparing the
general outputs of these programs will provide an idea of their
performance in different \textit{Trichoderma} genomes In addition to
these basic outputs, analysis will also be performed for the
following: distribution of gene lengths, intersection of gene calls,
smallRNAs and repetetive regions, shared gene content with a close
fungal relative. Analysis for these results can be performed through
simple shell scripting with grep and other unix tools, although
processing through Python might provide results that are easier to
reproduce with proper programming. Having one script with several
modules that can be rerun at will would be easier to handle than
multiple shell scripts. This thinking for processing will be applied
to subsequent sections of this as well.

\subsection{Distribution of Gene Lengths}

One important aspect of gene finding tools to consider is the
distribution of gene lengths predicted by each individual
tool. Certain tools, such as GeneMark are based on pre-defined models,
which may limit the length of predicted genes, while tools such as
Braker2, which incorporate RNAseq data, may predict a wider
distribution of gene lengths depending on the input dataset
used. Regardless, the ability of a gene finding tool to predict a
wider range of gene lengths can be usefull if users are looking for
short or larger genes. To help determine whether or not these tools
find shorter genes, or small RNAs, the genomes of interest have been
annotated using Infernal along with the Rfam database to identify
small RNAs as a ground truth. These annotation results will also be
included with results from other annotation processes further down the
line. Again, these results can be produced with a Python script. The
resulting data could then be used as input to violin plots for each
genome and set of tools considered in this analysis process. Violin
plots should provide a good visualization of gene lengths as well as
the number of genes found with specific lengths. Means could also be
compared staatistically for genomes and the mutliple tools considered
as well.

Analysis of gene lengths was performed using a Python script. Combined
predicted CDS sequences for each predicted gene were used as input for
the total gene length. CDS sequences predicted by Braker2, were
directly available in the output directories when the program was
run. CDS sequences from GeneMark required extraction of the CDS
sequences from the genome FASTA files. This process was performed
using the gffread tool from the Cufflinks package. Predicted CDS
sequences were loaded into Python using Biopython's SeqIO
package. Sequence lengths were then placed in a list and analyzed
using a combination of pandas and numpy. A log10 transformation was
applied to the sequence lengths as the original distribution was
heavily skewed due to long outlier CDS sequences. After
transformation, the CDS length distribution appears as a normal
distribution, although there are interesting troughs that occur in
several of the peaks for several of the genomes considered. These
troughs did not appear in a comparable Yeast reference dataset,
although the log transformed data still appears to be a normal
distribution.

\subsection{Intersection of Gene Calls, smallRNAs and Repetitive Regions}

Annotation of all three features in the title are important in
assessing the ability of gene finding tools. Even more important, is
the potential for for overlap between gene calls and small RNAS as
well as repetitive regions the genomes. As discussed in the previous
subsection, the distribution of gene lengths predicted by a gene
finding tool can be an important metric for users. Overlapping
predicted genes from tools alongside the output from Infernal and the
Rfam database may provide insight into whether or not these gene
finding tools are able to predict RNAs of very short length. In
addition to small RNAs, repetitive regions in \textit{Trichoderma}
genomes hold potential for recombination and gene content, although
the inherent nature of these repetitive regions (low nucleotide
diveristy) suggests that gene content should be low, based on the
nucleotides required for start and stop codons. Analysis of these
intersections can be perfomed via bedtools or through biopython (I
believe). Again, having all processing steps included in one script as
separate functions that can be called at whim will make further
processing easier if changes need to be made.

\subsection{Methodology for Indetifying Overlapping Features}

To analyze the results from multiple gene-finding tools, we must first
define two conceptual topics.

Definition of a Feature: First, a feature, in the context of this
research, is any item contained within a Genomic Feature Formated file
(GFF). Each feature contains a contig ID, a start position, end
position, and a feature ID based on the information from the GFF
file. These features will be used in the process of identifying
regions, or overlapping features from prediction tools.

Definition of a Region: Secondly, a region is defined as any overlap
between features relative to the reference sequence being
considered. To identify regions, every feature from each GFF file
being considered, is sorted by start position. For clarification, the
start position is based on the left most position in the GFF file,
regardless of the strand that the feature is predicted on. Once the
features have been sorted by left position, the sorted features are
iterated over to identify regions, as long as the left position of the
next feature is consistent with the left position , within the left
and right position, or equal to the right position of the current
region. One drawback to this approach is that ideally, identification
of overlap types based on Allen's interval algebra would be performed
at this point. However, the methodology of initially sorting features
based on left position somewhat prevents this process from
happening. This implementation requires further processing of
identified regions late on the process, simplifying the all to all
comparison that would occurr if all features were considered at once.

\subsection{Shared Gene Content with Yeast}

While considering novel gene calls can be useful, comparing those
calls to a well-studied close relative can provide a rudimentary
validation of the calls as a ground truth. In this case, a comparison
to Yeast will be made. The agreed upon number for successful
gene-finding as compared to Yeast is roughly 80-85 percent of gene
content. This will confirm that at least most of a closely related
fungal genome's content is predicted and shared by the gene calls for
\textit{Trichoderma}. Results for this processing can be produced with
a simple BLAST search and apropriate cutoff values (i.e. query
coverage, percent nucelotide identity, E-score, etc.). Other tools for
evaluation will certainly be considered, although I will need to look
into this process further.

\subsection{Comparative Genomics}

With the data produced by this research, it is possible to perform som
commparative genomics (time permitted), mostly related to the
assemblies generated during this work along with the RefSeq genomes
included from NCBI. Mummer is a potential tool to use for all to all
genome alignments, although there may be difficulty in the ordering of
contigs/scaffolds/chromosomes when performing thes alignments. This
work is not necessarily required but would be interesting from a
biological perspective to identify rearrangements, inversions and
such.
